---
title: "Lab 04 - Data Type Coercion"
author: "Nyla Williams"
date: "3//6/2024"
output: html_document
---

### Load packages and data

```{r load-packages message = FALSE}
library(tidyverse)
library(openintro)
```

## Exercises

### Exercise 1

```{r}
c(1, 1L, "C")
```

```{r}
## [1] 1
```

```{r}
## [1] 1
```

The output ([1] 1)  display the number 1. they are integers therefore they are numerical data types

```{r}
## [1] "C"
```


```{r}
# [1] "1" "1" "C"
```
creates a character vector by combining all three components. R allows various data types that are joined into a vector to conform to one unified type—in this example, character.


### Exercise 2


```{r}
c(1:3, 5)
```

there are only numericals here
```{r}
##[1] 1 2 3
```
.
```{r}
# [1] 5
```
when uncommented, it is still a numerical vector


### Exercise 3

```{r}
c(3, FALSE, T)
```
there are two logicals and a double.
```{r}
## [1] 3
```


```{r}
## [1] FALSE
```

```{r}
# [1] TRUE
```
logical data type

### Exercise 4

```{r}
c(TRUE, as.numeric(TRUE))
```
there is a double and a logical
```{r}
## [1] TRUE
```

```{r}
## [1] 1
```


### Exercise 5

```{r}
 c(NA, FALSE, as.logical(1))
```

There are all logicals
```{r}
## [1] NA
```

```{r}
## [1] FALSE
```

```{r}
## [1] TRUE
```

logicial 




### Exercise 6

```{r}
c(1L / 0, "A")
```
In this code chunk we see a double, a numerical, and a character
```{r}
## [1] 1
```

```{r}
## [1] 0
```

```{r}
## [1] Inf
```

```{r}
## [1] "A"
```

it is a character

### Exercise 7

```{r}
glimpse(earthquakes)
```


### Exercise 8

magnitude and depth

### Exercise 9

if we coerced "year" it may lead to some visualization issues; when plotting the year on the x-axis,it will be ordered incorrectly 

### Exercise 10

```{r}
earthquakes %>%
  mutate(major = case_when(
    richter > 7.5 ~ TRUE,
    TRUE ~ FALSE
  )) %>%
  mutate(major = as.numeric(major)) %>%
  filter(major == 1)
```
 From what we see here, there are 29 earthquakes in this data set.
